Coupling finite and boundary element methods to solve the Poisson-Boltzmann equation for electrostatics in molecular solvation

### 2022-10-21 methods.tex
Moving BEM-BEM section
I have moved the BEM formulation to the results because I do not know how appendices work in this journal. I have made some changes to make readable the method section so please check it.

There are regularized versions of Equation \eqref{eq:pbe} \cite{LuZhouHolstMcCammon2008,LeeGengZhao2021} which are widely used to numerically solve the Poisson-Boltzmann equation with finite element or finite difference methods.
However, here we use the standard formulation in Equation \eqref{eq:pbe}, as it offers more flexibility when dealing with, for example, variable permitivitties.

@@

\subsection*{\sffamily \large Numerical solution of the Poisson-Boltzmann equation}

\subsubsection*{\sffamily \normalsize BEM-BEM coupling}

The boundary element method (BEM) is a standard tool for the numerical solution of the Poisson-Boltzmann equation in molecular electrostatics.~\cite{ZauharMorgan1985, Shaw1985} This was implemented in numerous codes, such as AFMPB,~\cite{LuETal2006} TABI,~\cite{GengKrasny2013} PyGBe,~\cite{CooperBardhanBarba2014,cooper2016pygbe} and more recently, with Bempp-cl.~\cite{search2022towards}

There are several possible boundary integral formulations of Equation \eqref{eq:pbe}.~\cite{search2022towards} The simplest form was presented by Yoon ºand Lenhoff in 1991,~\cite{YoonLenhoff1990} known as the \emph{direct} formulation, however, the resulting system is usually ill-conditioned. Here, we use the better conditioned formulation presented by Lu and co-workers,~\cite{LuETal2006,LuETal2009,debuhr2016dashmm} which reads
%
\begin{align}\label{eq:lu}
    & \tfrac{\phi_2}{2}\left(1+\tfrac{\epsilon_1}{\epsilon_2}\right) - \left(K_Y^\Gamma - \tfrac{\epsilon_1}{\epsilon_2}K_L^\Gamma\right)\phi_2 + \left(V_Y^\Gamma - V_L^\Gamma\right)\lambda_2 = \sum_{k=0}^{N_q}  \frac{q_k}{4\pi\epsilon_2|\mathbf{x}_{\Gamma} - \mathbf{x}_k|}
     \nonumber \\
    &\tfrac{\epsilon_1}{\epsilon_2}\left(W_Y^\Gamma - W_L^\Gamma\right)\phi_2 +  \tfrac{\lambda_2}{2}\left(1+\tfrac{\epsilon_1}{\epsilon_2}\right) + \left(\tfrac{\epsilon_1}{\epsilon_2}K_Y^{\prime\Gamma} - K_L^{\prime\Gamma}\right)\lambda_2 = \sum_{k=0}^{N_q}  \frac{\partial}{\partial\mathbf{n}_\mathbf{x}}\left(\frac{q_k}{4\pi\epsilon_2|\mathbf{x}_{\Gamma} - \mathbf{x}_k|}\right)
\end{align}
%
In Equation \eqref{eq:lu}, $\phi_1 = \phi_1(\mathbf{x}_\Gamma)$ and $\phi_2 = \phi_2(\mathbf{x}_\Gamma)$ are the potentials on $\Gamma$ as we approach from $\Omega_1$ and $\Omega_2$, respectively. For convenience, we call the normal derivative on the interface ({\it ie.} the Neumann trace) $\lambda=\left(\frac{\partial}{\partial \mathbf{n}}  \phi_{\Gamma}  \right)$. $V$, $K$, $K^{\prime}$, and $W$ are the single-layer, double-layer, adjoint double-layer, and hypersingular operators for the Laplace (subscript $L$) and Yukawa (subscript $Y$, also known as modified Helmholtz) kernels. These are defined as
%
\begin{align}\label{eq:single_double}
V_i^\Gamma \varphi (\mathbf{x}) &= \oint_\Gamma g_i(\mathbf{x},\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}'\nonumber\\
K_i^\Gamma \varphi (\mathbf{x}) &= \oint_\Gamma \frac{\partial g_i}{\partial\mathbf{n}'}(\mathbf{x},\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}',\nonumber\\
K^{\prime\Gamma}_i\varphi (\mathbf{x}) &= \oint_\Gamma \frac{g_i}{\partial\mathbf{n}}(\mathbf{x}_\Gamma,\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}'\nonumber\\
W^\Gamma_i\varphi (\mathbf{x}) &= - \oint_\Gamma \frac{\partial^2 g_i}{\partial\mathbf{n}'\partial\mathbf{n}}(\mathbf{x}_\Gamma,\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}'
\end{align}
%
where $i \in \{L,Y\}$, $\varphi(\mathbf{x})$ can be any distribution over $\Gamma$, and
%
\begin{align}\label{eq:green_func}
g_L(\mathbf{x},\mathbf{x}')=\frac{1}{4\pi|\mathbf{x}-\mathbf{x}'|} \nonumber \\
g_Y(\mathbf{x},\mathbf{x}')=\frac{e^{-\kappa|\mathbf{x}-\mathbf{x}'|}}{4\pi|\mathbf{x}-\mathbf{x}'|}
\end{align}
%
are the corresponding free-space Green's functions.

%Applying Green's second identity to Equation \eqref{eq:pbe} we get 
%
%\begin{align} \label{eq:volume_potential}
%\phi_{1}(\mathbf{x}_{\Omega_1})+ K_{L}^{\Omega_1}\phi_1 -  V_{L}^{\Omega_1} \lambda_2^1 & = \tfrac{1}{\epsilon_1} \sum_{k=0}^{N_q}  \frac{q_k}{4\pi|\mathbf{x}_{\Omega_1} - \mathbf{x}_k|}  \quad \text{on $\Omega_1$,} \nonumber \\
%\phi_{2}(\mathbf{x}_{\Omega_2}) - K_{Y}^{\Omega_2}\phi_2 + \tfrac{\epsilon_1}{\epsilon_2} V_{Y}^{\Omega_2} \lambda_2& = 0 \quad \text{on $\Omega_2$,}
%\end{align}
%
%where $\mathbf{x}_\Omega$ is evaluated anywhere in the correspoding domain, except $\Gamma$, and $\phi_1 = \phi_1(\mathbf{x}_\Gamma)$ and $\phi_2 = \phi_2(\mathbf{x}_\Gamma)$ are the potentials on $\Gamma$ as we approach from $\Omega_1$ and $\Omega_2$, respectively. For convenience, we call the normal derivative on the interface ({\it ie.} the Neumann trace) $\lambda=\left(\frac{\partial}{\partial \mathbf{n}}  \phi_{\Gamma}  \right)$. $K$ and $V$ are the double- and single-layer potentials for the Laplace (subscript $L$) and Yukawa (subscript $Y$, also known as modified Helmholtz) kernels
%
%\begin{align}\label{eq:single_double}
%V^D \varphi (\mathbf{x}) = \oint_\Gamma g_i(\mathbf{x},\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}'\nonumber\\
%K^D \varphi (\mathbf{x}) = \oint_\Gamma \frac{\partial g_i}{\partial\mathbf{n}'}(\mathbf{x},\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}',\nonumber\\
%\end{align}
%
%where $D \in \{\Omega_1, \Omega_2, \Gamma\}$, $i \in \{L,Y\}$, $\varphi(\mathbf{x})$ can be any distribution over $\Gamma$, and 
%
%\begin{align}\label{eq:green_func}
%g_L(\mathbf{x},\mathbf{x}')=\frac{1}{4\pi|\mathbf{x}-\mathbf{x}'|} \nonumber \\
%g_Y(\mathbf{x},\mathbf{x}')=\frac{e^{-\kappa|\mathbf{x}-\mathbf{x}'|}}{4\pi|\mathbf{x}-\mathbf{x}'|}
%\end{align}
%
%are the free-space Green's function of the Laplace and linearized Poisson-Boltzmann equations, respectively.

%If we take the limit of Eq. \eqref{eq:volume_potential} as $\mathbf{x}_{\Omega}$ approaches $\Gamma$, we get
%
%\begin{align} \label{eq:direct}
%\tfrac{\phi_1}{2}+ K_{L}^{\Gamma}\phi_1 -  V_{L}^{\Gamma} \lambda_2^1 & = \tfrac{1}{\epsilon_1} \sum_{k=0}^{N_q}  \frac{q_k}{4\pi|\mathbf{x}_{\Gamma} - \mathbf{x}_k|} \nonumber \\
%\tfrac{\phi_2}{2} - K_{Y}^{\Gamma}\phi_2 +  \tfrac{\epsilon_1}{\epsilon_2} V_{Y}^{\Gamma} \lambda_2 & = 0
%\end{align}
%
%The {\it direct} formulation\cite{YoonLenhoff1990} couples the two expressions of Eq. \eqref{eq:direct} on $\Gamma$, considering the interface conditions. Even though this formulation is poorly conditioned with respect to the mesh size, it can still model large systems when it is preconditioned appropriately.\cite{AltmanBardhanWhiteTidor09,MartinezETal2019,wang2021high} 
%The formulation presented by Lu\cite{LuETal2006} takes the normal derivative of \eqref{eq:direct} to couple both $\phi$ and $\partial\phi/\partial\mathbf{n}$ as
%
%\begin{align}\label{eq:lu}
%    & \tfrac{\phi_2}{2}\left(1+\tfrac{\epsilon_1}{\epsilon_2}\right) - \left(K_Y^\Gamma - \tfrac{\epsilon_1}{\epsilon_2}K_L^\Gamma\right)\phi_2 + \left(V_Y^\Gamma - V_L^\Gamma\right)\lambda_2 = \sum_{k=0}^{N_q}  \frac{q_k}{4\pi\epsilon_2|\mathbf{x}_{\Gamma} - \mathbf{x}_k|}
%     \nonumber \\
%    &\tfrac{\epsilon_1}{\epsilon_2}\left(W_Y^\Gamma - W_L^\Gamma\right)\phi_2 +  \tfrac{\lambda_2}{2}\left(1+\tfrac{\epsilon_1}{\epsilon_2}\right) + \left(\tfrac{\epsilon_1}{\epsilon_2}K_Y^{\prime\Gamma} - K_L^{\prime\Gamma}\right)\lambda_2 = \sum_{k=0}^{N_q}  \frac{\partial}{\partial\mathbf{n}_\mathbf{x}}\left(\frac{q_k}{4\pi\epsilon_2|\mathbf{x}_{\Gamma} - \mathbf{x}_k|}\right)
%\end{align}
%
%The two new integral operators that appear are defined as 
%
%\begin{align}
%\label{eq:adj_hyp}
%K^{\prime\Gamma}_i\varphi (\mathbf{x}) = \oint_\Gamma \frac{g_i}{\partial\mathbf{n}}(\mathbf{x}_\Gamma,\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}'\nonumber\\
%W^\Gamma_i\varphi (\mathbf{x}) = - \oint_\Gamma \frac{\partial^2 g_i}{\partial\mathbf{n}'\partial\mathbf{n}}(\mathbf{x}_\Gamma,\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}'
%\end{align}
%where $i \in \{L,Y\}$, 
%and are called the {\it  adjoint double layer} and {\it hypersingular} operators, respectively. Note that the potential in Eq. \eqref{eq:lu} is evaluated on the exterior.

In weak form, Eq. \eqref{eq:lu} becomes
%
\begin{align}\label{eq:lu_disc}
&\left< \left(1+\tfrac{\epsilon_1}{\epsilon_2}\right) \tfrac{\phi_{2}}{2} - \left(K_Y^\Gamma - \tfrac{\epsilon_1}{\epsilon_2}K_L^\Gamma\right)\phi_{2}, v \right>_{\Gamma} + \left< \left(V_Y^\Gamma - V_L^\Gamma\right) \lambda_2, v \right>_\Gamma = \left< \sum_{k=0}^{N_q}  f_k, v \right>_{\Gamma} \nonumber \\
&\left< \tfrac{\epsilon_1}{\epsilon_2}\left(W_Y^\Gamma - W_L^\Gamma\right)\phi_{2}, \zeta \right>_\Gamma + \left<  \left(1+\tfrac{\epsilon_1}{\epsilon_2}\right) \tfrac{\lambda_2}{2} + \left(\tfrac{\epsilon_1}{\epsilon_2}K_Y^{\prime\Gamma} - K_L^{\prime\Gamma}\right)\lambda_2, \zeta\right>_\Gamma = \left< \sum_{k=0}^{N_q}  \frac{\partial}{\partial\mathbf{n}_\mathbf{x}} f_k, \zeta \right>_{\Gamma}
\end{align}
%
where $v, \zeta$ are test functions, $\left<\varphi,v\right>_\Gamma = \int_\Gamma \varphi(\mathbf{x})v(\mathbf{x})d\mathbf{x}$ %and $\left(\varphi,v\right)_\Omega = \int_\Omega \varphi(\mathbf{x})v(\mathbf{x})d\mathbf{x}$ are 
is the inner products on the surface% and domain, respectively
and $f_k := \frac{q_k}{4\pi\epsilon_2|\mathbf{x}_{\Gamma} - \mathbf{x}_k|}$.

We introduce below the matrix formulation. Let $\vec{\phi}_2 := [\phi_2^1, \dots, \phi_2^k]^T$ be the vector of canonical basis
functions of $W_{h}^{k}$ and  $\vec{\lambda}_2 := [\lambda_2^1, \dots, \lambda_2^l]^T$ - the vector of canonical basis
functions of $\Lambda_{h}^{l}$. We define the following matrices associated with the corresponding bilinear forms
\begin{align*}
\widetilde{K}_{\alpha \beta} = \left<\left(K_Y^\Gamma - \tfrac{\epsilon_1}{\epsilon_2}K_L^\Gamma\right) \phi^{\alpha}_2, \ \lambda^{\beta}_2 \right>_{\Gamma}
 &,&
\widetilde{V}_{\alpha \beta} = \left<\left(V_Y^\Gamma - V_L^\Gamma\right) \lambda^{\alpha}_2, \ \lambda^{\beta}_2 \right>_{\Gamma}, \\
\widetilde{W}_{\alpha \beta} = \left<\left(W_Y^\Gamma - W_L^\Gamma\right) \phi^{\alpha}_2, \ \phi^{\beta}_2 \right>_{\Gamma}
 &,&
\widetilde{K'}_{\alpha \beta} = \left<\left(\tfrac{\epsilon_1}{\epsilon_2}K_Y^{\prime\Gamma} - K_L^{\prime\Gamma}\right) \lambda^{\alpha}_2, \ \phi^{\beta}_2 \right>_{\Gamma},
\end{align*}
and vectors associated with the corresponding linear forms
\begin{align*}
\vec{f}_{\beta} = \left< \sum_{k=0}^{N_q}  \frac{q_k}{4\pi\epsilon_2|\mathbf{x}_{\Gamma} - \mathbf{x}_k|},   \phi^{\beta}_1  \right>_{\Gamma} &,& \vec{\partial f}_{\beta} = \left< \sum_{k=0}^{N_q}  \frac{\partial}{\partial\mathbf{n}_\mathbf{x}}\left(\frac{q_k}{4\pi\epsilon_2|\mathbf{x}_{\Gamma} - \mathbf{x}_k|}\right),   \phi^{\beta}_1  \right>_{\Gamma}.
\end{align*}

Using the above definition the discrete problem~\eqref{eq:fem_bem_disctrete} can be written in the following matrix form
\begin{align*}
\begin{bmatrix}
 \tfrac12 \left(1+\tfrac{\epsilon_1}{\epsilon_2}\right) I - \widetilde{K}  & \widetilde{V}  \\  
 \tfrac{\epsilon_1}{\epsilon_2} \widetilde{W} & \tfrac12 \left(1+\tfrac{\epsilon_1}{\epsilon_2}\right) I + \widetilde{K'}
\end{bmatrix}
\begin{bmatrix}
\vec{\phi}_2 \\
\vec{\lambda}_2 
\end{bmatrix}
= 
\begin{bmatrix}
\vec{f} \\  
\vec{\partial f}
\end{bmatrix}.
\end{align*}
%

Having computed the electrostatic potential with Eq. \eqref{eq:lu_disc}, we obtain the reaction potential ($\phi_r$) in Eq. \eqref{eq:dG} by subtracting out the Coulombic component out of $\phi_1$. This gives\cite{CooperBardhanBarba2014}
%
\begin{align} \label{eq:phi_reac}
\phi_{r}(\mathbf{r}) = - K_{L}^{\mathbf{r}}(\phi_1) + V_{L}^{\mathbf{r}} \left(  \lambda_1  \right),
\end{align}
%
where $\mathbf{r}\in\Omega_1$.
This is valid as long as $\epsilon_1$ is a constant. We use Eq.~\eqref{eq:phi_reac} for both BEM-BEM and FEM-BEM coupling simulations with constant $\epsilon_1$.

\subsection*{\sffamily \normalsize FEM-BEM coupling}

BEM is a popular alternative to solve the Poisson-Boltzmann equation numerically because it reduces the dimension of the problem by using the boundary integral formulation. Unfortunately, this advantage comes at a cost: it requires a fundamental solution to be applied. There are many linear problems for which fundamental solutions are not known, for example, this is the case for the Poisson-Boltzmann equation with a heterogenous permittivity inside the molecule
%
   \begin{align} \label{eq:pbe_vp}
\nabla \cdot \left(\epsilon_1(\mathbf{x}) \nabla \phi_1(\mathbf{x})\right) &= \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k) \quad  \mathbf{x} \in \Omega_1\nonumber\\
\left(\nabla^2 - \kappa^2\right)\phi_2(\mathbf{x})  &= 0 \quad\mathbf{x}\in\Omega_2\nonumber\\
\phi_1(\mathbf{x})  = \phi_2(\mathbf{x}),  &\quad \epsilon_1(\mathbf{x})\frac{\partial\phi_1}{\partial\mathbf{n}}(\mathbf{x})  = \epsilon_2(\mathbf{x})\frac{\partial\phi_2}{\partial\mathbf{n}}(\mathbf{x})  \quad \mathbf{x}\in \Gamma. 
\end{align}
%
For such case, a finite element method (FEM) is more suitable.

@@

   We start with the variational formulation of the internal problem. Applying integration by parts for first equation of~\eqref{eq:pbe_vp} for every $v \in H_0^1(\Omega_1)$, we have
% 
\begin{equation}
\label{eq:fem}
 \left( \epsilon_1 \nabla \phi_1, \nabla v \right)_{\Omega_1}  -  \left<  \epsilon_1\partial_n \phi_1, v \right>_\Gamma =  \left( \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k),  v\right)_{\Omega_1},
\end{equation}
%
where $v$ is a test function, and $\left<\varphi,v\right>_\Gamma = \int_\Gamma \varphi(\mathbf{x})v(\mathbf{x})d\mathbf{x}$ and $\left(\varphi,v\right)_{\Omega_1} = \int_{\Omega_1} \varphi(\mathbf{x})v(\mathbf{x})d\mathbf{x}$ are the inner products on the surface and domain, respectively.
For the external problem with BEM, we use the direct formulation from Eq.~\eqref{eq:direct} to obtain
\begin{align*}
\tfrac{\phi_1}{2} - K_{Y}^{\Gamma}(\phi_1) + \tfrac{\epsilon_1}{\epsilon_2}V_{Y}^{\Gamma}  \left(\lambda_1 \right) & = 0.
\end{align*}
%
Then, the coupling problem can be written as
%
\begin{center}
  \textit{Find $ \phi_1 \in H^1(\Omega_1)$ and $\lambda_2 \in H^{-\frac{1}{2}}(\Gamma)$ such that for all $v \in H^1(\Omega_1)$ and $\zeta \in H^{-\frac{1}{2}}(\Gamma)$}
\begin{equation} 
\label{eq:standard_fem_bem}
 \left\{
 \begin{array}{rcl}
 \left(  \epsilon_1 \nabla \phi_1, \nabla v \right)_{\Omega_1}  - \left< \epsilon_1 \lambda_2, v \right>_\Gamma &=&   \left(  \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k),  v \right)_{\Omega_1} \\[3mm] 
  \left< \left(\tfrac{1}{2} I - K_{Y}^{\Gamma}\right) \phi_1, \zeta \right>_\Gamma + \tfrac{\epsilon_1}{\epsilon_2} \left< V_{Y}^{\Gamma} \lambda_2, \zeta \right>_\Gamma &=&0.
  \end{array}
  \right.
\end{equation}
\end{center}

@@

This can also be written in matrix form. Let $\vec{\phi}_1 := [\phi_1^1, \dots, \phi_1^j]^T$ be the vector of canonical basis
functions of the finite element space $V_{h}^{j}$, %$\vec{\phi}_2 := [\phi_2^1, \dots, \phi_2^k]^T$ - the vector of canonical basis functions of $W_{h}^{k}$,  
$\vec{\lambda}_2 := [\lambda_2^1, \dots, \lambda_2^l]^T$ --- the vector of canonical basis
functions of $\Lambda_{h}^{l}$. % and $\vec{\widetilde{\phi}} := [\widetilde{\phi}^1, \dots, \widetilde{\phi}^m]^T$ be the vector of canonical basis functions of  $M_{h}^{m}$
% $T:\Omega_1 \rightarrow \Gamma$ and $E:\Gamma \rightarrow \Omega_1$ trace and extendable operators respectively.
We define the following matrices associated with the corresponding bilinear forms
\begin{align*}
A_{\alpha \beta} = \left(\nabla \phi_1, \nabla \phi_1^{\beta} \right)_{\Omega_1}  
&,&
M_{\alpha \beta} = \left< \widetilde{\phi}^{\alpha}, \ \lambda^{\beta}_2 \right>_{\Gamma}, \\
K_{\alpha \beta} = \left<K_{Y}^{\Gamma} \phi^{\alpha}_1, \ \lambda^{\beta}_2 \right>_{\Gamma}
 &,&
V_{\alpha \beta} = \left<V_{Y}^{\Gamma} \lambda^{\alpha}_2, \ \lambda^{\beta}_2 \right>_{\Gamma}, 
\end{align*}
and vector associated with the corresponding linear form
\begin{equation*}
\vec{f}_{\beta} := \left(  \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k),  \phi_1^{\beta} \right)_{\Omega_1}.
\end{equation*}
Using the above definition the discrete problem in Eq.~\eqref{eq:standard_fem_bem} can be written in the following matrix form
\begin{align*}
\begin{bmatrix}
\epsilon_1 A &  - \epsilon_1 M^T \\  
\left(\tfrac12 I + K \right) &  \tfrac{\epsilon_1}{\epsilon_2} V 
\end{bmatrix}
\begin{bmatrix}
\vec{\phi}_1 \\  
\vec{\lambda}_2
\end{bmatrix}
= 
\begin{bmatrix}
\vec{f} \\  
0
\end{bmatrix}.
\end{align*}

>>>>

Regularized versions of Equation \eqref{eq:pbe} \cite{LuZhouHolstMcCammon2008,LeeGengZhao2021} are widely used to numerically solve the Poisson-Boltzmann equation with finite element or finite difference methods.
However, here we use the standard formulation in Equation \eqref{eq:pbe}, as it offers more flexibility when dealing with, for example, variable permitivitties.

@@

\subsection*{\sffamily \large Novel numerical solution of the Poisson-Boltzmann equation}

%\subsection*{\sffamily \normalsize FEM-BEM coupling}

The boundary element method (BEM) is a standard tool for the numerical solution of the Poisson-Boltzmann equation in molecular electrostatics.~\cite{ZauharMorgan1985, Shaw1985} This was implemented in numerous codes, such as AFMPB,~\cite{LuETal2006} TABI,~\cite{GengKrasny2013} PyGBe,~\cite{CooperBardhanBarba2014,cooper2016pygbe} and more recently, with Bempp-cl.~\cite{search2022towards}. Its popularity comes because it reduces the dimension of the problem by using the boundary integral formulation. Unfortunately, this advantage comes at a cost: it requires a fundamental solution to be applied. There are many linear problems for which fundamental solutions are not known, for example, this is the case for the Poisson-Boltzmann equation with a heterogenous permittivity inside the molecule
%
   \begin{align} \label{eq:pbe_vp}
\nabla \cdot \left(\epsilon_1(\mathbf{x}) \nabla \phi_1(\mathbf{x})\right) &= \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k) \quad  \mathbf{x} \in \Omega_1\nonumber\\
\left(\nabla^2 - \kappa^2\right)\phi_2(\mathbf{x})  &= 0 \quad\mathbf{x}\in\Omega_2\nonumber\\
\phi_1(\mathbf{x})  = \phi_2(\mathbf{x}),  &\quad \epsilon_1(\mathbf{x})\frac{\partial\phi_1}{\partial\mathbf{n}}(\mathbf{x})  = \epsilon_2(\mathbf{x})\frac{\partial\phi_2}{\partial\mathbf{n}}(\mathbf{x})  \quad \mathbf{x}\in \Gamma. 
\end{align}
%
A finite element method (FEM) is more suitable for such case.

@@

   We start with the variational formulation of the internal problem. Applying integration by parts for first equation of~\eqref{eq:pbe_vp} for every $v \in H_0^1(\Omega_1)$, we have
% 
\begin{equation}
\label{eq:fem}
 \left( \epsilon_1 \nabla \phi_1, \nabla v \right)_{\Omega_1}  -  \left<  \epsilon_1\partial_n \phi_1, v \right>_\Gamma =  \left( \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k),  v\right)_{\Omega_1},
\end{equation}
%
where $v$ is a test function, and $\left<\varphi,v\right>_\Gamma = \int_\Gamma \varphi(\mathbf{x})v(\mathbf{x})d\mathbf{x}$ and $\left(\varphi,v\right)_{\Omega_1} = \int_{\Omega_1} \varphi(\mathbf{x})v(\mathbf{x})d\mathbf{x}$ are the inner products on the surface and domain, respectively.
For the external problem with BEM, we define the Dirichlet trace~\cite{MR2361676} 
\begin{align*}
% \label{eq:trace_map}
\gamma:  H^1(\Omega_2) &\rightarrow H^{\frac{1}{2}}(\Gamma) & \gamma f(x) & := \lim_{\Omega_2 \ni y \rightarrow x \in \Gamma}  f(y),
\end{align*}
and we use the direct formulation from the second equation of Eq.~\eqref{eq:pbe} to obtain
\begin{align*}
\tfrac{\phi_1}{2} - K_{Y}^{\Gamma}(\gamma \phi_1) + \tfrac{\epsilon_1}{\epsilon_2}V_{Y}^{\Gamma}  \left(\lambda_1 \right) & = 0.
\end{align*}
$V$ and $K$ are the single-layer and double-layer operators for the Laplace (subscript $L$) and Yukawa (subscript $Y$, also known as modified Helmholtz) kernels. These are defined as
%
\begin{align}\label{eq:single_double}
V_i^\Gamma \varphi (\mathbf{x}) &= \oint_\Gamma g_i(\mathbf{x},\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}'\nonumber\\
K_i^\Gamma \varphi (\mathbf{x}) &= \oint_\Gamma \frac{\partial g_i}{\partial\mathbf{n}'}(\mathbf{x},\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}',\nonumber\\
%K^{\prime\Gamma}_i\varphi (\mathbf{x}) &= \oint_\Gamma \frac{g_i}{\partial\mathbf{n}}(\mathbf{x}_\Gamma,\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}'\nonumber\\
%W^\Gamma_i\varphi (\mathbf{x}) &= - \oint_\Gamma \frac{\partial^2 g_i}{\partial\mathbf{n}'\partial\mathbf{n}}(\mathbf{x}_\Gamma,\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}'
\end{align}
%
where $i \in \{L,Y\}$, $\varphi(\mathbf{x})$ can be any distribution over $\Gamma$, and
%
\begin{align}\label{eq:green_func}
g_L(\mathbf{x},\mathbf{x}')=\frac{1}{4\pi|\mathbf{x}-\mathbf{x}'|} \nonumber \\
g_Y(\mathbf{x},\mathbf{x}')=\frac{e^{-\kappa|\mathbf{x}-\mathbf{x}'|}}{4\pi|\mathbf{x}-\mathbf{x}'|}
\end{align}
%
are the corresponding free-space Green's functions.

Then, the coupling problem can be written as
%
\begin{center}
  \textit{Find $ \phi_1 \in H^1(\Omega_1)$ and $\lambda_2 \in H^{-\frac{1}{2}}(\Gamma)$ such that for all $v \in H^1(\Omega_1)$ and $\zeta \in H^{-\frac{1}{2}}(\Gamma)$}
\begin{equation} 
\label{eq:standard_fem_bem}
 \left\{
 \begin{array}{rcl}
 \left(  \epsilon_1 \nabla \phi_1, \nabla v \right)_{\Omega_1}  - \left< \epsilon_1 \lambda_2, v \right>_\Gamma &=&   \left(  \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k),  v \right)_{\Omega_1} \\[3mm] 
  \left< \left(\tfrac{1}{2} I - K_{Y}^{\Gamma}\right) \gamma \phi_1, \zeta \right>_\Gamma + \tfrac{\epsilon_1}{\epsilon_2} \left< V_{Y}^{\Gamma} \lambda_2, \zeta \right>_\Gamma &=&0.
  \end{array}
  \right.
\end{equation}
\end{center}

@@

This can also be written in matrix form. Let $\vec{\phi}_1 := [\phi_1^1, \dots, \phi_1^j]^T$ be the vector of canonical basis
functions of the finite element space $V_{h}^{j}$, %$\vec{\phi}_2 := [\phi_2^1, \dots, \phi_2^k]^T$ - the vector of canonical basis functions of $W_{h}^{k}$,  
$\vec{\lambda}_2 := [\lambda_2^1, \dots, \lambda_2^l]^T$ --- the vector of canonical basis
functions of $\Lambda_{h}^{l}$. % and $\vec{\widetilde{\phi}} := [\widetilde{\phi}^1, \dots, \widetilde{\phi}^m]^T$ be the vector of canonical basis functions of  $M_{h}^{m}$
% $T:\Omega_1 \rightarrow \Gamma$ and $E:\Gamma \rightarrow \Omega_1$ trace and extendable operators respectively.
We define the following matrices associated with the corresponding bilinear forms
\begin{align*}
A_{\alpha \beta} = \left(\nabla \phi_1^{\alpha}, \nabla \phi_1^{\beta} \right)_{\Omega_1}  
&,&
M_{\alpha \beta} = \left< \widetilde{\phi}^{\alpha}, \ \lambda^{\beta}_2 \right>_{\Gamma}, \\
K_{\alpha \beta} = \left<K_{Y}^{\Gamma} \gamma \phi^{\alpha}_1, \ \lambda^{\beta}_2 \right>_{\Gamma}
 &,&
V_{\alpha \beta} = \left<V_{Y}^{\Gamma} \lambda^{\alpha}_2, \ \lambda^{\beta}_2 \right>_{\Gamma}, 
\end{align*}
and vector associated with the corresponding linear form
\begin{equation*}
\vec{f}_{\beta} := \left(  \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k),  \phi_1^{\beta} \right)_{\Omega_1}.
\end{equation*}
Using the above definition the discrete problem in Eq.~\eqref{eq:standard_fem_bem} can be written in the following matrix form
\begin{align*}
\begin{bmatrix}
\epsilon_1 A &  - \epsilon_1 M^T \\  
\left(\tfrac12 I + K \right) \gamma &  \tfrac{\epsilon_1}{\epsilon_2} V 
\end{bmatrix}
\begin{bmatrix}
\vec{\phi}_1 \\  
\vec{\lambda}_2
\end{bmatrix}
= 
\begin{bmatrix}
\vec{f} \\  
0
\end{bmatrix}.
\end{align*}

Having computed the electrostatic potential with Eq. \eqref{eq:standard_fem_bem}, we obtain the reaction potential ($\phi_r$) in Eq. \eqref{eq:dG} by subtracting out the Coulombic component out of $\gamma \phi_1$. This gives\cite{CooperBardhanBarba2014}
%
\begin{align} \label{eq:phi_reac}
\phi_{r}(\mathbf{r}) = - K_{L}^{\mathbf{r}}( \gamma \phi_1) + V_{L}^{\mathbf{r}} \left(  \lambda_1  \right),
\end{align}
%
where $\mathbf{r}\in\Omega_1$.
This is valid as long as $\epsilon_1$ is a constant. We use Eq.~\eqref{eq:phi_reac} for both BEM-BEM and FEM-BEM coupling simulations with constant $\epsilon_1$.
